\documentclass[conference]{IEEEtran}

\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}
\usepackage{listings}
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{tikz}
\usetikzlibrary{arrows.meta,positioning,shapes.geometric,fit,calc,backgrounds}
\usepackage{pgfplots}
\pgfplotsset{compat=1.18}
\usepackage{hyperref}
\usepackage{balance}

% Code listing style
\lstset{
  basicstyle=\footnotesize\ttfamily,
  breaklines=true,
  frame=single,
  captionpos=b
}

\begin{document}

\title{SHADOW: Seamless Handoff And Zero-Downtime\\Orchestrated Workload Migration\\for Stateful Microservices}

\author{
\IEEEauthorblockN{Hai Dinh-Tuan}
\IEEEauthorblockA{
\textit{Technische Universit\"at Berlin} \\
Berlin, Germany \\
hai.dinh-tuan@tu-berlin.de}
}

\maketitle

%% CONDENSED: original two-paragraph abstract
% Live migration of stateful microservices in Kubernetes remains challenging due to
% the tight coupling between container state, pod identity, and orchestrator
% constraints. Our prior work introduced the Message-based Stateful Microservice
% Migration (MS2M) framework, which reconstructs service state through message
% replay rather than low-level memory transfer, and demonstrated its integration
% with Kubernetes using Forensic Container Checkpointing. However, the initial
% implementation relied on an external Python-based migration manager and suffered
% from two performance bottlenecks: a restore phase dominated by StatefulSet
% identity constraints forcing a sequential stop-recreate cycle, and a checkpoint
% transfer phase consuming several seconds for the OCI registry round-trip.
%
% This paper presents SHADOW ... [full original text preserved above]
%% CONDENSED: previous single-paragraph abstract
% This paper presents SHADOW (...) a Kubernetes-native framework for live
% migration of stateful microservices that builds on the MS2M approach.
% SHADOW makes three contributions: (1) Kubernetes Operator with CRD and
% idempotent state machine; (2) ShadowPod strategy eliminating ~38s restore
% bottleneck; (3) direct node-to-node transfer. Evaluation: 92% restore
% reduction, 0ms downtime, 73% total time reduction, zero message loss.
\begin{abstract}
Migrating stateful microservices in Kubernetes requires careful state
management because in-memory state is lost when a container restarts.
For StatefulSet-managed workloads, the problem is compounded by identity
constraints that prohibit two pods with the same ordinal from running
simultaneously, forcing a sequential stop-recreate cycle with a median
38.5\,s of service downtime. This paper presents SHADOW
(\textbf{S}eamless \textbf{H}andoff \textbf{A}nd Zero-\textbf{D}owntime
\textbf{O}rchestrated \textbf{W}orkload Migration), a Kubernetes-native
framework that implements the Message-based Stateful Microservice Migration
(MS2M) approach as a Kubernetes Operator. SHADOW introduces the
\textit{ShadowPod} strategy, where a shadow pod is created from a CRIU
checkpoint image on the target node while the source pod continues serving
traffic, allowing concurrent operation during message replay. For
StatefulSet workloads, an identity swap procedure with the ExchangeFence
mechanism re-checkpoints the shadow pod, creates a StatefulSet-owned
replacement, and drains both message queues to guarantee zero message loss
during the handoff. An evaluation on a bare-metal Kubernetes cluster with
280 migration runs across four configurations and seven message rates
(10--120\,msg/s) shows that, compared to the sequential baseline on the same
StatefulSet workload, the ShadowPod strategy reduces the restore phase by up
to 92\%, eliminates service downtime entirely, and reduces total migration
time by up to 77\%, with zero message loss across all 280 runs.
\end{abstract}

\begin{IEEEkeywords}
microservices, live migration, Kubernetes, container checkpointing,
CRIU, stateful services, operator pattern
\end{IEEEkeywords}


%% ============================================================================
\section{Introduction}
\label{sec:introduction}
%% ============================================================================

The increasing adoption of microservice architectures in cloud-native and
cloud/edge environments has introduced new challenges around service
mobility. While stateless services can be rescheduled across cluster
nodes, stateful microservices that maintain in-memory state for
performance-critical operations require careful state management during
migration. Traditional migration techniques, originally developed for virtual
machines, treat service state as an opaque memory block and rely on pre-copy or
post-copy mechanisms that introduce additional complexity and network
overhead~\cite{ma2019live, tazzioli2023postcopy}.

The Message-based Stateful Microservice Migration (MS2M)
framework~\cite{dinhtuan2022ms2m} addresses this challenge by using the
application's own messaging infrastructure to reconstruct state on the target
node. Rather than copying raw memory, MS2M checkpoints the running container,
restores it on a target node, and replays buffered messages to synchronize
state, reducing downtime to the brief checkpoint creation phase. A subsequent
extension~\cite{dinhtuan2025k8s} integrated MS2M into Kubernetes using Forensic
Container Checkpointing (FCC)~\cite{k8sfcc}, introduced a threshold-based cutoff
mechanism for bounded replay times, and adapted the procedure for StatefulSet
workloads using a Sequential (stop-then-recreate) strategy.

However, the Kubernetes integration in~\cite{dinhtuan2025k8s} exhibited two
performance bottlenecks. First, the restore phase for StatefulSet
pods required a median of 38.5 seconds due to Kubernetes' identity constraints,
which prohibit two pods with the same StatefulSet-managed hostname from running
simultaneously. This forced a sequential stop-recreate cycle, during which the
service was completely unavailable. Second, the checkpoint transfer phase
consumed a median of 5.4 seconds for a 27\,MB checkpoint, as the process involved
building an OCI-compliant container image, pushing it to a container registry,
and pulling it on the target node. Additionally, the migration was orchestrated
by an an out-of-band agent (\texttt{Migration Manager}) that communicated with the cluster via
\texttt{kubectl}, operating outside Kubernetes' declarative resource model.

This paper presents SHADOW\footnote{The SHADOW implementation is publicly
available at \url{https://github.com/haidinhtuan/shadow}} (Seamless Handoff
And Zero-Downtime Orchestrated Workload Migration), a Kubernetes-native
framework that addresses these limitations through three contributions:

%% CONDENSED: detailed contribution descriptions (moved detail to System Design)
% \item \textbf{Kubernetes Operator.} SHADOW implements MS2M as a Kubernetes
%   Operator using a Custom Resource Definition (\texttt{StatefulMigration}) and a
%   controller-runtime-based reconciler. The operator encodes the five-phase
%   migration procedure as an idempotent state machine, replacing the external
%   Python script with a Kubernetes-native control loop that integrates with the
%   platform's desired-state model, RBAC policies, and observability
%   infrastructure.
%
% \item \textbf{ShadowPod Migration Strategy.} A migration strategy is introduced
%   that creates a \textit{shadow pod} on the target node from the checkpoint
%   image while the source pod remains running. Both pods operate concurrently
%   during the message replay phase, eliminating the sequential restore
%   bottleneck. Crucially, this strategy is not limited to
%   Deployment-managed workloads: for StatefulSet-managed pods, the shadow pod is
%   created with a distinct name (e.g., \texttt{consumer-0-shadow}) outside the
%   StatefulSet's ownership, carrying the same application labels for Service
%   routing. During finalization, the StatefulSet is scaled down by one replica to
%   remove the original pod, while the shadow pod continues serving traffic. This
%   avoids the identity conflict without requiring a change in workload type.
%
% \item \textbf{Direct Node-to-Node Transfer.} An \texttt{ms2m-agent}
%   DaemonSet is deployed on worker nodes that receives checkpoint archives
%   directly from the source node via HTTP, builds the OCI image locally, and
%   loads it into the container runtime's image store---bypassing the container
%   registry entirely.
\begin{enumerate}
\item \textit{Kubernetes Operator:} a Custom Resource Definition
  (\texttt{StatefulMigration}) and an idempotent state machine reconciler that
  replaces the Migration Manager with a Kubernetes-native control loop.

\item \textit{ShadowPod Migration Strategy:} a shadow pod is created on the
  target node from the checkpoint image while the source pod continues serving,
  allowing concurrent operation during message replay. This works for
  both Deployment- and StatefulSet-managed workloads.

\item \textit{Identity Swap via ExchangeFence:} after ShadowPod migration
  on a StatefulSet, the shadow pod is orphaned outside the StatefulSet's
  ownership. SHADOW provides an identity swap procedure that re-checkpoints
  the shadow pod, creates a StatefulSet-owned replacement, and uses the
  ExchangeFence mechanism (unbinding, draining queues, and rebinding) to guarantee zero message loss during this handoff.
\end{enumerate}

%% CONDENSED: scoping and evaluation summary (already in abstract)
% SHADOW targets event-driven microservices whose in-memory state is derived from
% processed messages rather than persistent disk storage. The framework is
% evaluated on a bare-metal Kubernetes cluster across four configurations and
% 280 total migration runs.

%% CONDENSED: paper organization paragraph (filler)
% The remainder of this paper is organized as follows. Section~\ref{sec:background}
% provides background on MS2M and Kubernetes integration challenges.
% Section~\ref{sec:design} presents the SHADOW design, including the operator
% architecture and both optimization strategies. Section~\ref{sec:evaluation}
% details the experimental evaluation and discusses the results.
% Section~\ref{sec:related} surveys related work, and Section~\ref{sec:conclusion}
% concludes the paper.


%% ============================================================================
\section{Background and Prior Work}
\label{sec:background}
%% ============================================================================

\subsection{The MS2M Framework}
\label{sec:ms2m}

The Message-based Stateful Microservice Migration (MS2M)
framework~\cite{dinhtuan2022ms2m} was designed for event-driven microservices
that derive their state from processed message streams. Unlike traditional
migration approaches that treat service state as an opaque memory block, MS2M
uses a property of message-driven architectures: \textit{service state
is the deterministic result of the messages a service has processed}. This
makes state reconstruction possible through message replay rather than low-level
memory transfer.

%% CONDENSED: original enumerate list for five-phase procedure
% The framework defines a five-phase migration procedure coordinated by a
% Migration Manager:
%
% \begin{enumerate}
% \item \textbf{Checkpoint Creation:} The source container is briefly paused using
%   CRIU (Checkpoint/Restore in Userspace)~\cite{criu} to create a process
%   checkpoint. A secondary message queue is created to buffer incoming messages
%   from this point forward. The source is resumed immediately after
%   checkpointing.
%
% \item \textbf{Checkpoint Transfer:} The checkpoint archive is transferred from
%   the source host to the target host. The source service continues processing
%   live traffic during this phase.
%
% \item \textbf{Service Restoration:} The checkpoint is restored on the target
%   host, creating a new service instance with the exact process state captured
%   at checkpoint time.
%
% \item \textbf{Message Replay:} The restored instance processes all buffered
%   messages from the secondary queue, synchronizing its state with the source's
%   state at the time of checkpointing.
%
% \item \textbf{Finalization:} The target instance switches from the secondary
%   queue to the primary queue for live traffic. The source instance is
%   terminated, and the secondary queue is deleted.
% \end{enumerate}
The framework defines a five-phase migration procedure coordinated by a
Migration Manager: (1)~\textbf{Checkpoint Creation} -- the source container is
briefly paused using CRIU~\cite{criu} to create a process checkpoint while a
secondary queue buffers incoming messages; (2)~\textbf{Checkpoint Transfer} --
the archive is sent to the target host while the source continues serving;
(3)~\textbf{Service Restoration} -- the checkpoint is restored on the target;
(4)~\textbf{Message Replay} -- the restored instance processes buffered messages
from the secondary queue; and (5)~\textbf{Finalization} -- the target switches
to the primary queue and the source is terminated.

%% CONDENSED: expanded key assumption paragraph
% A key assumption of the MS2M framework is that the target workload's state
% is fully determined by the messages it has processed. During the replay phase,
% the source and target instances consume from \textit{separate} queues (primary
% and secondary, respectively), so no message is consumed twice from the same
% queue. However, the same logical messages are present in both queues due to the
% fanout exchange. If the service produces downstream side effects (e.g., database
% writes, outbound events), these may be duplicated during replay. Applications
% using MS2M must therefore either suppress side effects during the replay phase
% (detectable via the \texttt{START\_REPLAY}/\texttt{END\_REPLAY} control
% messages) or ensure that downstream operations are idempotent.
A key assumption is that service state is fully determined by processed
messages. During replay, source and target consume from separate queues, so no
message is consumed twice. Applications must suppress side effects during replay
(detectable via \texttt{START\_REPLAY}/\texttt{END\_REPLAY} control messages) or
ensure idempotent downstream operations.

%% CONDENSED: original proof-of-concept details
% The original proof-of-concept~\cite{dinhtuan2022ms2m} was built on Podman with
% raw CRIU and direct host-to-host transfer via rsync, orchestrated by an
% imperative Python script. It demonstrated a 19.92\% reduction in service
% downtime compared to traditional stop-and-copy migration.
The original proof-of-concept~\cite{dinhtuan2022ms2m} demonstrated a 19.92\%
reduction in service downtime compared to traditional stop-and-copy migration.

\subsection{Kubernetes Integration}
\label{sec:k8s-integration}

%% CONDENSED: expanded Kubernetes integration first paragraph
% The integration of MS2M into Kubernetes~\cite{dinhtuan2025k8s} required
% adapting the framework to the platform's declarative, desired-state resource
% model. The original three-actor architecture (Source Host, Target Host,
% Migration Manager) was extended to a four-actor model that includes the
% Kubernetes API Server as an intermediary for all cluster operations.
%
% A critical technical foundation for this integration is \textit{Forensic
% Container Checkpointing} (FCC)~\cite{k8sfcc}, an experimental Kubernetes
% feature (since v1.25) that extends the kubelet API to create CRIU checkpoints
% of running containers. FCC enables checkpointing without requiring direct
% access to the container runtime or CRIU binary on worker nodes. The checkpoint
% transfer was implemented by packaging the checkpoint archive as an OCI-compliant
% container image and pushing it to a container registry, from which the target
% node pulls during pod creation.
%
% The prior work also identified two key challenges:
The integration of MS2M into Kubernetes~\cite{dinhtuan2025k8s} required
adapting the framework to the platform's declarative, desired-state resource
model. A critical technical foundation is \textit{Forensic Container
Checkpointing} (FCC)~\cite{k8sfcc}, an experimental Kubernetes feature (since
v1.25) that extends the kubelet API to create CRIU checkpoints of running
containers without requiring direct access to the container runtime. Checkpoint
transfer was implemented by packaging the archive as an OCI-compliant container
image and pushing it to a container registry, from which the target node pulls
during pod creation. That work identified two key challenges:

\textbf{Unbounded Replay Time.} When the incoming message rate approaches the
service's processing capacity, the replay queue grows indefinitely, making
migration time unpredictable. A \textit{Threshold-Based Cutoff Mechanism} was
introduced to address this: the source service is terminated after a calculated
duration $T_{\text{cutoff}} \leq \frac{T_{\text{replay\_max}} \cdot
\mu_{\text{target}}}{\lambda}$, bounding the number of messages to replay and
guaranteeing migration completion within a specified time window.

\textbf{StatefulSet Constraints.} StatefulSet pods possess stable, unique network
identities that Kubernetes enforces by prohibiting two pods with the same ordinal
index from running simultaneously. This prevents the concurrent source-target
operation that MS2M relies on, forcing a sequential stop-recreate cycle
(\textit{Sequential strategy}) that dominated migration time at a median of
38.5 seconds. The prior work treated this as an inherent limitation of StatefulSet
workloads.

\subsection{Kubernetes Operators}
\label{sec:operators}

A Kubernetes Operator~\cite{operatorpattern} is a design pattern that combines
a Custom Resource Definition (CRD) with a dedicated controller to manage
application-specific operational knowledge. The controller implements a
\textit{reconcile loop} that continuously compares the desired state (expressed
in the custom resource's \texttt{spec}) with the actual cluster state and takes
corrective actions to converge. This declarative model provides automatic retry
on failure, leader election for high availability, integration with Kubernetes
RBAC and audit logging, and native observability through resource status fields
and Kubernetes events.


%% ============================================================================
\section{System Design}
\label{sec:design}
%% ============================================================================

This section presents the architecture of SHADOW and its two optimization
strategies. Figure~\ref{fig:architecture} provides an overview of the system
components.

\begin{figure*}[t]
\centering
\begin{tikzpicture}[
  >=Stealth,
  % Styles
  comp/.style={draw, rounded corners=2pt, thick, minimum height=8mm,
               minimum width=24mm, font=\footnotesize, align=center, fill=white},
  grp/.style={draw, thick, dashed, rounded corners=5pt, inner sep=12pt},
  glabel/.style={font=\small\bfseries},
  annot/.style={font=\scriptsize, fill=white, inner sep=1.5pt},
  flow/.style={->, thick},
  % Helper for side-by-side positioning
  node distance=0.5cm and 1.2cm
]

% --- 1. Control Plane ---
\node[comp, fill=blue!10] (operator) at (1.5, 0) {MS2M Operator\\[-1pt]{\scriptsize (Reconciler)}};
\node[comp, fill=blue!10] (apiserver) at (7.5, 0) {API Server};
\draw[<->, thick] (operator) -- node[annot, above] {watch/update StatefulMigration CR} (apiserver);

% --- 2. Worker 1 (Source) ---
\node[comp, fill=green!12, minimum width=20mm] (kubelet1) at (0.5, -3) {Kubelet\\[-1pt]{\scriptsize CRI-O + CRIU}};
\node[comp, fill=yellow!20, minimum width=20mm] (agent1) at (2.8, -3) {ms2m-agent};
\node[comp, fill=orange!15] (srcpod) at (1.65, -4.8) {Source Pod\\[-1pt]{\scriptsize (Consumer)}};

% --- 3. Worker 2 (Target) --- mirrored: agent inside, kubelet outside
\node[comp, fill=yellow!20, minimum width=20mm] (agent2) at (6.2, -3) {ms2m-agent};
\node[comp, fill=green!12, minimum width=20mm] (kubelet2) at (8.5, -3) {Kubelet\\[-1pt]{\scriptsize CRI-O + CRIU}};
\node[comp, fill=orange!15] (tgtpod) at (7.35, -4.8) {Shadow Pod\\[-1pt]{\scriptsize (Restored)}};

% --- 4. Message Broker ---
\node[comp, fill=red!10] (primaryq) at (2.5, -7.5) {Primary Queue};
\node[comp, fill=red!10] (replayq) at (6.5, -7.5) {Replay Queue};
\node[comp, fill=red!5, minimum width=45mm] (rabbitmq) at (4.5, -8.6) {RabbitMQ};

% --- Background Boxes ---
\begin{scope}[on background layer]
  \node[grp, fill=blue!5, fit=(operator)(apiserver), inner ysep=15pt] (cpbox) {};
  \node[grp, fill=green!5, fit=(kubelet1)(agent1)(srcpod)] (w1box) {};
  \node[grp, fill=green!5, fit=(kubelet2)(agent2)(tgtpod)] (w2box) {};
  \node[grp, fill=red!5, fit=(primaryq)(replayq)(rabbitmq), inner ysep=10pt] (mqbox) {};
\end{scope}

% Labels
\node[glabel, anchor=south] at (cpbox.north) {Control Plane};
\node[glabel, anchor=south] at (w1box.north) {Worker 1 (Source)};
\node[glabel, anchor=south] at (w2box.north) {Worker 2 (Target)};
\node[glabel, anchor=south] at (mqbox.north) {Message Broker};

% --- 5. Flow Arrows ---

% Step 1: To Kubelet
\draw[flow, blue!70] (operator.south) -- ++(0,-0.6) -|
  node[annot, pos=0.25, right] {\textbf{1.} Checkpoint} (kubelet1.north);

% Step 2: Agent to Agent (Direct Transfer) - now a clear path
\draw[flow, red!60, dashed, line width=1.1pt] (agent1.east) --
  node[annot, above] {\textbf{2.} Direct transfer} (agent2.west);

% Step 3: To Kubelet 2
\draw[flow, blue!70] (apiserver.south) -- ++(0,-0.6) -|
  node[annot, pos=0.25, left] {\textbf{3.} Shadow pod} (kubelet2.north);

% Step 4: Finalize
\node[annot, font=\scriptsize\itshape, text=black!80] at (4.5, -5.6)
  {\textbf{4.} Finalize: switch traffic, remove source};

% Queue Interactions
\draw[flow, red!60] (srcpod.south) -- ++(0,-0.4) -|
  node[annot, near end, left] {consume} (primaryq.north);
\draw[flow, red!60] (replayq.north) -- ++(0,0.4) -|
  node[annot, near start, right] {replay} (tgtpod.south);

% Broker internal connections
\draw (rabbitmq.north -| primaryq.south) -- (primaryq.south);
\draw (rabbitmq.north -| replayq.south) -- (replayq.south);

\end{tikzpicture}
\caption{Architecture of the SHADOW framework. The reconciler watches
StatefulMigration custom resources and orchestrates migration through the API
server: checkpoint creation via the kubelet API~(\textbf{1}), checkpoint
transfer between ms2m-agent instances or via registry~(\textbf{2}), shadow pod
creation on the target node~(\textbf{3}), and traffic switchover during
finalization~(\textbf{4}).}
\label{fig:architecture}
\end{figure*}

\subsection{The SHADOW Operator}
\label{sec:operator}

SHADOW implements the MS2M migration procedure as a Kubernetes Operator using the
\texttt{controller-runtime} framework~\cite{controllerruntime}. The operator
manages a custom resource, \texttt{StatefulMigration}, whose specification
declares the migration intent and whose status reflects the current progress.

\subsubsection{Custom Resource Definition}
\label{sec:crd}

The \texttt{StatefulMigration} CRD (API group: \texttt{migration.ms2m.io/v1alpha1})
captures the migration intent through the following key specification fields:

\begin{itemize}
\item \texttt{sourcePod}: The name of the pod to migrate.
\item \texttt{targetNode}: The destination worker node.
\item \texttt{migrationStrategy}: \texttt{Sequential} or \texttt{ShadowPod}.
  Auto-detected from the pod's owner reference chain if not specified.
\item \texttt{transferMode}: \texttt{Registry} (default) or \texttt{Direct}
  (node-to-node via ms2m-agent).
\item \texttt{messageQueueConfig}: RabbitMQ connection details, queue name, and
  exchange name for the replay mechanism.
\item \texttt{replayCutoffSeconds}: Maximum duration for the cutoff mechanism.
\end{itemize}

The resource status tracks the current \texttt{phase}, per-phase durations
(\texttt{phaseTimings}), the checkpoint identifier, and metadata cached from
the source pod (labels, container specs, owner references).
%% CONDENSED: metadata caching rationale (original listed "owning StatefulSet or
%% Deployment name, original replica count" instead of "owner references")
% This metadata caching is necessary because in
% Sequential strategy, the source pod is deleted before the target pod is created,
% so its labels and container specifications must be preserved in the CRD status
% during the Pending phase.

\subsubsection{State Machine Reconciler}
\label{sec:statemachine}

The reconciler implements the MS2M five-phase procedure as an idempotent state
machine with seven phases, illustrated in Figure~\ref{fig:statemachine}:
\texttt{Pending}, \texttt{Checkpointing},
\texttt{Transferring}, \texttt{Restoring}, \texttt{Replaying},
\texttt{Finalizing}, and \texttt{Completed} (or \texttt{Failed}). Each
reconciliation invocation dispatches to the handler for the current phase, which
either advances to the next phase or re-queues for a later attempt.

Phase handlers are \textit{idempotent}: re-entering a phase after a transient
failure does not corrupt migration state. The reconciler implements
\textit{phase chaining} (executing synchronously-completed phases within a
single API call) and \textit{exponential polling backoff} for long-running
phases, reducing API server load. Wall-clock durations for each phase are
recorded in \texttt{status.phaseTimings} for built-in instrumentation.
%% CONDENSED: expanded idempotency and optimization details
% Each handler first checks whether its phase-specific work has already been
% completed before initiating new actions. This property is essential for
% production reliability, as the reconcile loop may be triggered multiple times
% for the same phase due to watch events, cache synchronization, or controller
% restarts.
% Phase chaining: when a phase handler completes synchronously (returns an
% immediate requeue), the reconciler re-fetches the resource and enters the next
% handler within the same API call, avoiding a work queue round-trip.
% Exponential polling backoff: long-running asynchronous phases (Transferring,
% Restoring, Replaying) use adaptive requeue intervals that increase from 1s to
% 2s to 5s based on elapsed wall-clock time.

\begin{figure}[t]
\centering
\begin{tikzpicture}[
  >=Stealth,
  state/.style={draw, rounded corners=3pt, thick, minimum width=24mm,
    minimum height=6mm, font=\footnotesize},
  trans/.style={font=\scriptsize, fill=white, inner sep=1pt},
  arr/.style={->, thick},
  fail/.style={->, dashed, thick, red!50!black}
]
\node[state, fill=gray!12]  (P) at (0,0)    {Pending};
\node[state, fill=blue!10]  (C) at (0,-1.1) {Checkpointing};
\node[state, fill=blue!10]  (T) at (0,-2.2) {Transferring};
\node[state, fill=blue!10]  (R) at (0,-3.3) {Restoring};
\node[state, fill=blue!10]  (Y) at (0,-4.4) {Replaying};
\node[state, fill=blue!10]  (F) at (0,-5.5) {Finalizing};
\node[state, fill=green!12] (D) at (0,-6.6) {Completed};
\node[state, fill=red!10]   (X) at (3.2,-3.3) {Failed};

% Happy path
\draw[arr] (P) -- node[trans, right] {pod validated} (C);
\draw[arr] (C) -- node[trans, right] {CRIU checkpoint} (T);
\draw[arr] (T) -- node[trans, right] {image ready} (R);
\draw[arr] (R) -- node[trans, right] {pod running} (Y);
\draw[arr] (Y) -- node[trans, right] {queue drained} (F);
\draw[arr] (F) -- node[trans, right] {traffic switched} (D);

% Failure transitions
\draw[fail] (C.east) to[out=0,in=155] (X);
\draw[fail] (T.east) to[out=0,in=175] (X);
\draw[fail] (R.east) -- (X);
\draw[fail] (Y.east) to[out=0,in=185] (X);
\draw[fail] (F.east) to[out=0,in=205] (X);
\node[font=\scriptsize\itshape, text=red!50!black, anchor=west]
  at (X.east) {~on error};
\end{tikzpicture}
\caption{State machine of the StatefulMigration reconciler. The reconcile loop
dispatches to the handler for the current phase, advancing on success or
transitioning to Failed on error. Each handler is idempotent for safe retry.}
\label{fig:statemachine}
\end{figure}

%% CONDENSED: Strategy Detection subsubsection
% \subsubsection{Strategy Detection}
% \label{sec:strategy-detection}
% When \texttt{migrationStrategy} is not explicitly specified, the operator
% auto-detects the appropriate strategy by traversing the source pod's owner
% reference chain. If the chain leads to a StatefulSet, the Sequential strategy
% is selected; if it leads to a ReplicaSet (and transitively a Deployment), the
% ShadowPod strategy is selected. This auto-detection can be overridden: setting
% \texttt{migrationStrategy: ShadowPod} explicitly on a StatefulSet-managed pod
% activates the ShadowPod strategy with StatefulSet-specific finalization logic,
% as described in Section~\ref{sec:shadowpod-statefulset}.
When \texttt{migrationStrategy} is not specified, the operator auto-detects it
from the pod's owner reference chain (StatefulSet $\to$ Sequential, Deployment
$\to$ ShadowPod), but this can be overridden explicitly.


\subsection{ShadowPod Migration Strategy}
\label{sec:shadowpod}

The ShadowPod strategy removes the sequential restore bottleneck by allowing
source and target pods to operate concurrently during the replay phase.
Kubernetes Services route traffic based on \textit{label selectors}, while
StatefulSet identity constraints are enforced through \textit{owner references}.
A shadow pod that carries the same application
labels but is not owned by the StatefulSet can coexist with the original pod
and receive Service traffic without violating identity constraints.

\subsubsection{Core Procedure}
\label{sec:shadowpod-core}

Figure~\ref{fig:timeline} illustrates the key difference between the Sequential
and ShadowPod strategies. In the Sequential approach, the source pod must be
terminated before the target pod can be created, causing a service gap. In the
ShadowPod approach, the shadow pod is created and starts serving alongside the
source, providing an uninterrupted handoff.

\begin{figure}[t]
\centering
\begin{tikzpicture}[scale=0.9, every node/.style={scale=0.9},
  phase/.style={draw, thick, minimum height=5mm, font=\scriptsize,
    inner sep=2pt, anchor=west},
  label/.style={font=\scriptsize\bfseries, anchor=east},
  gap/.style={draw, thick, fill=red!20, minimum height=5mm,
    font=\scriptsize, inner sep=2pt, anchor=west},
]
% --- Sequential Timeline ---
\node[label] at (-0.2,1.8) {Sequential};
\node[label] at (-0.2,1.2) {Source:};
\node[phase, fill=blue!15, minimum width=8mm] at (0,1.2) {Ckpt};
\node[phase, fill=yellow!20, minimum width=16mm] at (0.8,1.2) {Transfer};
\node[gap, minimum width=32mm] at (2.4,1.2) {\textcolor{red!70!black}{DOWN (38.5\,s)}};

\node[label] at (-0.2,0.6) {Target:};
\node[phase, fill=green!20, minimum width=32mm] at (2.4,0.6) {Restore};
\node[phase, fill=purple!15, minimum width=20mm] at (5.6,0.6) {Replay};
\node[phase, fill=orange!15, minimum width=4mm] at (7.6,0.6) {F};

% --- ShadowPod Timeline ---
\node[label] at (-0.2,-0.4) {ShadowPod};
\node[label] at (-0.2,-1.0) {Source:};
\node[phase, fill=blue!15, minimum width=8mm] at (0,-1.0) {Ckpt};
\node[phase, fill=green!15, minimum width=56mm] (srcrun)
  at (0.8,-1.0) {Running (serves traffic)};

\node[label] at (-0.2,-1.6) {Shadow:};
\node[phase, fill=yellow!20, minimum width=16mm] at (0.8,-1.6) {Transfer};
\node[phase, fill=green!20, minimum width=6mm] at (2.4,-1.6) {R};
\node[phase, fill=purple!15, minimum width=32mm] at (3.0,-1.6) {Replay};
\node[phase, fill=orange!15, minimum width=4mm] at (6.2,-1.6) {F};
\node[phase, fill=green!15, minimum width=12mm] at (6.6,-1.6) {Serves};

\node[font=\scriptsize\bfseries, text=green!50!black] at (4.5,-2.3)
  {0\,ms downtime};

% Separator
\draw[dashed, gray] (-0.2,0.15) -- (8.2,0.15);
\end{tikzpicture}
\caption{Timeline comparison of Sequential vs.\ ShadowPod migration. In
Sequential, service is unavailable during the 38.5\,s median restore phase. In
ShadowPod, the source continues serving while the shadow pod restores and
replays, achieving zero downtime.}
\label{fig:timeline}
\end{figure}

The ShadowPod strategy modifies the Restoring and Finalizing phases of the
migration procedure. During the \textbf{Restoring} phase, instead of scaling
down a StatefulSet and recreating its pod, the operator creates a new pod
named \texttt{<source>-shadow} on the target node. This shadow pod uses the
checkpoint container image with \texttt{imagePullPolicy: Never} (since the
image is already present in the target node's local image store). The shadow
pod carries the same application labels as the source pod, so it can
receive traffic via the Kubernetes Service immediately upon becoming ready.
Both the source pod and the shadow pod run concurrently during this phase
and throughout the subsequent replay phase.

%% CONDENSED: expanded readiness/routing discussion
% Because the shadow pod is restored from a CRIU checkpoint, its HTTP health
% endpoint becomes available immediately after restore---the CRIU process image
% includes the running HTTP server thread. Kubernetes marks the shadow pod as
% Ready once the health probe succeeds, adding it to the Service's endpoint list.
% During the replay phase, both pods serve traffic: the source pod with current
% state, and the shadow pod with progressively-updating state (starting from
% checkpoint time and converging to current as replay messages are processed).
% This is acceptable for the target workload class (event-driven microservices)
% because reads from the shadow pod return valid, consistent state---albeit
% briefly stale until replay completes. The Kubernetes Service load-balances
% across both endpoints, and the source pod ensures that clients always have
% access to a fully up-to-date instance throughout the migration.
Because the shadow pod is restored from a CRIU checkpoint, its health endpoint
is available immediately after restore. Kubernetes adds it to the Service's
endpoint list, so both pods serve traffic during replay. The source provides
fully up-to-date state while the shadow pod's state converges as replay
progresses.

\subsubsection{Deployment-Managed Workloads}
\label{sec:shadowpod-deployment}

For pods owned by a Deployment, the \textbf{Finalizing} phase executes the
following handoff sequence:

\begin{enumerate}
\item Send \texttt{END\_REPLAY} to the shadow pod, causing it to switch
  from the replay queue to the primary queue.
\item Delete the source pod.
\item Patch the owning Deployment's pod template with a \texttt{nodeAffinity}
  rule targeting the destination node, ensuring that future pods scheduled by
  the Deployment controller land on the target node.
\item Delete the replay queue and close the broker connection.
\end{enumerate}

The Deployment controller manages the desired replica count and will
schedule replacement pods according to the updated affinity rule.

\subsubsection{StatefulSet-Managed Workloads}
\label{sec:shadowpod-statefulset}

For pods owned by a StatefulSet, the finalization must account for the fact
that direct pod deletion would cause the StatefulSet controller to immediately
recreate the pod (restoring the replica count). Instead, the \textbf{Finalizing}
phase:

\begin{enumerate}
\item Sends \texttt{END\_REPLAY} to the shadow pod.
\item Scales the owning StatefulSet down by one replica. The StatefulSet
  controller removes the highest-ordinal pod (i.e., the original source pod),
  while the shadow pod (which is not owned by the StatefulSet) continues
  serving traffic.
\item Deletes the replay queue and closes the broker connection.
\end{enumerate}

This procedure preserves the key property of ShadowPod migration: the source
remains operational during the entire replay phase, and the handoff is
instantaneous from the Service's perspective since the shadow pod is already
receiving traffic via label-based routing.

\subsubsection{Trade-offs}
\label{sec:shadowpod-tradeoffs}

For StatefulSets, the ShadowPod strategy results in the workload being served
by a standalone shadow pod outside StatefulSet ownership. StatefulSet guarantees
(ordered scaling, PVC attachments) do not apply. SHADOW therefore targets
in-memory stateful services whose state is derived from message processing, not
persistent disk storage. For Deployment-managed workloads, no such trade-off
exists.
%% CONDENSED: expanded trade-off discussion
% The StatefulSet's replica count has been decremented and the shadow pod exists
% outside its ownership. This means StatefulSet guarantees (ordered scaling,
% stable storage attachments) do not apply to the shadow pod. In particular,
% PersistentVolumeClaims (PVCs) bound to the StatefulSet are not automatically
% attached to the shadow pod. For workloads using ReadWriteOnce PVCs, the shadow
% pod cannot mount the volume while the source pod holds it, making this strategy
% inapplicable. For event-driven microservices whose routing is managed by the
% message broker rather than DNS-based pod identity, this trade-off is
% acceptable. To return to full StatefulSet management, the shadow pod can be
% deleted and the StatefulSet scaled back to its original replica count.
% For Deployment-managed workloads, the Deployment controller continues managing
% the desired replica count after migration, and the affinity patch ensures
% correct node placement.


\subsection{Identity Swap via ExchangeFence}
\label{sec:identity-swap}

The ShadowPod strategy for StatefulSets leaves the workload served by an orphaned shadow pod outside the StatefulSet's ownership. To restore full StatefulSet management, SHADOW provides an identity swap procedure during the Finalizing phase:

\begin{enumerate}
\item The shadow pod is re-checkpointed on the target node using a local CRIU dump.
\item The StatefulSet is scaled down by one replica, removing the original source pod.
\item A replacement pod with the original name (e.g., \texttt{consumer-0}) is created from the re-checkpoint image. The StatefulSet controller adopts this pod.
\item The shadow pod is terminated.
\end{enumerate}

During step~4, traffic must be handed off from the shadow pod to the replacement pod without message loss. A naive cutoff (terminating the shadow and starting the replacement) risks duplicate processing or lost messages if the handoff is not atomic. SHADOW addresses this with the \textit{ExchangeFence} mechanism:

\begin{enumerate}
\item A temporary buffer queue is bound to the message exchange.
\item The exchange is unbound from both the primary and swap queues, fencing new messages into the buffer.
\item Both the primary and swap queues are drained to zero.
\item The shadow pod is terminated, the primary queue is rebound to the exchange, and the buffer queue is drained into the replacement pod.
\end{enumerate}

The ExchangeFence provides a consistent cut: no messages are lost because the buffer queue captures all messages during the fence window, and no messages are duplicated because both source queues are fully drained before the handoff completes. This procedure adds 9--17\,s to the Finalizing phase (depending on message rate and queue depths) but restores full StatefulSet ownership of the migrated workload.


\subsection{Direct Node-to-Node Transfer}
\label{sec:direct-transfer}

The second optimization targets the container registry round-trip during
checkpoint transfer. SHADOW provides two alternatives to Job-based registry
transfer. First, when the \texttt{ms2m-agent} DaemonSet is present on the
source node, the controller delegates the registry push to the agent process,
avoiding the overhead of creating and scheduling a Kubernetes Job (agent-assisted
registry transfer). Second, the controller implements a direct transfer mode
through worker-to-worker communication that bypasses the registry entirely.
All four evaluation configurations use registry-based transfer; the direct
mode is not evaluated but is described here as a design contribution.

\subsubsection{The ms2m-agent DaemonSet}
\label{sec:agent}

The \texttt{ms2m-agent} is deployed as a Kubernetes DaemonSet on all worker nodes.
The agent exposes an HTTP endpoint (port 9443) that accepts checkpoint archives
and performs local image construction. The agent's pod specification includes
\texttt{hostPath} volume mounts for \texttt{/var/lib/kubelet/checkpoints}
(read-only, for accessing FCC checkpoint archives) and \texttt{/var/lib/ms2m}
(read-write, for temporary image construction).

\subsubsection{Transfer Procedure}
\label{sec:transfer-procedure}

When \texttt{transferMode: Direct} is specified, the Transferring phase operates
as follows:

\begin{enumerate}
\item The operator creates a transfer Job scheduled on the source node.
\item The Job reads the checkpoint archive from the kubelet's checkpoint
  directory.
\item The Job streams the archive as a multipart HTTP POST to the target
  node's ms2m-agent at
  \texttt{http://ms2m-agent.ms2m-system.svc:9443/checkpoint}.
\item The target agent writes the archive to disk, constructs an
  OCI-compliant container image (a single uncompressed layer with CRI-O
  checkpoint annotations), and loads it into the local image store using
  \texttt{skopeo copy}.
\item The agent responds with the local image reference, which the operator
  records in the migration status.
\end{enumerate}

%% CONDENSED: uncompressed layer rationale
% The OCI image is constructed with an uncompressed layer rather than a
% gzip-compressed one: on cluster-local networks, the CPU cost of compression
% outweighs the bandwidth savings, and the uncompressed image can be loaded
% faster by the container runtime.
%% CONDENSED: expanded registry fallback description
% This direct path eliminates the OCI image push and pull operations through the
% container registry. For the default \texttt{Registry} transfer mode, the
% existing procedure from~\cite{dinhtuan2025k8s} is preserved: the transfer Job
% builds the OCI image and pushes it to a configured registry, from which the
% target node pulls during pod creation.
The OCI image uses an uncompressed layer, as CPU compression cost outweighs
bandwidth savings on cluster-local networks. This direct path eliminates the
registry push/pull round-trip. The default \texttt{Registry} transfer mode
from~\cite{dinhtuan2025k8s} remains available as a fallback.

\subsubsection{CRIU Hostname Resolution}
\label{sec:hostname}

After CRIU restore, \texttt{gethostname()} returns the source pod's name (from
the UTS namespace snapshot), causing the control message protocol to listen on
the wrong queue. This is resolved by reading \texttt{/etc/hostname} (bind-mounted
by the runtime, reflecting the actual pod name) instead.
%% CONDENSED: expanded hostname resolution discussion
% During implementation, we identified a subtle interaction between CRIU
% checkpoint/restore and Kubernetes pod identity that affects the control message
% protocol. Since the MS2M control protocol routes START_REPLAY and END_REPLAY
% messages to ms2m.control.<pod-name>, the restored process would listen on the
% wrong control queue. This is a necessary adaptation for any CRIU-based
% migration system that uses pod-name-based routing in Kubernetes.


%% ============================================================================
\section{Evaluation}
\label{sec:evaluation}
%% ============================================================================

This section presents the experimental evaluation of SHADOW across four
migration configurations, comparing the baseline Sequential approach with the
ShadowPod strategy on both StatefulSet and Deployment workloads.

\subsection{Experimental Setup}
\label{sec:setup}

\subsubsection{Infrastructure}

The evaluation was conducted on a bare-metal Kubernetes cluster provisioned on
a European cloud provider, consisting of three dedicated servers:

\begin{itemize}
\item \textbf{Control plane:} 1 node running the Kubernetes API server,
  scheduler, controller manager, and the MS2M operator.
\item \textbf{Workers:} 2 nodes serving as alternating source and target for
  round-trip migrations.
\end{itemize}

Each server is equipped with 4 dedicated vCPUs, 8\,GB RAM, and 80\,GB SSD
storage. All nodes run Ubuntu 22.04 with Kubernetes v1.32, CRI-O as the
container runtime, and CRIU v4.0 compiled from source for checkpoint/restore
operations. An in-cluster container registry (deployed in the \texttt{registry}
namespace) is used for OCI checkpoint image transfer. RabbitMQ 3.13 is deployed
as a single-instance StatefulSet for message brokering.

\subsubsection{Workload}

The evaluation workload consists of two microservices:

\begin{itemize}
\item \textbf{Producer:} A Go Deployment that publishes messages to a RabbitMQ
  fanout exchange at a configurable rate. Each message contains a sequence
  number and timestamp.
\item \textbf{Consumer:} A single-replica workload (StatefulSet or Deployment,
  depending on configuration) implemented in Python. The consumer maintains an
  in-memory counter of processed messages and the last-seen sequence number as
  its stateful context. It exposes an HTTP health endpoint on port 8080 that
  reports the current processing state, serving both as a readiness probe and
  a downtime measurement target.
\end{itemize}

The consumer implements the MS2M control message protocol: it listens on
\texttt{ms2m.control.<hostname>} for \texttt{START\_REPLAY} and
\texttt{END\_REPLAY} messages. On \texttt{START\_REPLAY}, it switches
consumption from the primary queue to the replay queue specified in the message
payload. On \texttt{END\_REPLAY}, it switches back to the primary queue.

\subsubsection{Configurations}

Four migration configurations are evaluated, designed to isolate the effect
of the ShadowPod strategy and identity swap:

\begin{enumerate}
\item \textbf{statefulset-sequential:} The baseline configuration from prior
  work~\cite{dinhtuan2025k8s}. The consumer runs as a StatefulSet with the
  Sequential migration strategy. During the Restoring phase, the operator
  scales the StatefulSet to zero replicas, waits for the source pod to
  terminate, then creates the target pod with the same identity. Checkpoints
  are transferred via the in-cluster container registry.

\item \textbf{statefulset-shadowpod:} The consumer runs as a StatefulSet with
  the ShadowPod strategy (explicitly set via \texttt{migrationStrategy:
  ShadowPod} in the CR). The operator creates a shadow pod alongside the
  still-running source. During Finalizing, the StatefulSet is scaled down by
  one replica. Checkpoints are transferred via registry, isolating the effect
  of the ShadowPod strategy from the transfer mode optimization.

\item \textbf{deployment-registry:} The consumer runs as a Deployment with the
  ShadowPod strategy. The operator creates a shadow pod alongside the source
  and patches the Deployment's \texttt{nodeAffinity} during finalization.
  Checkpoints are transferred via registry.

\item \textbf{statefulset-shadowpod-swap:} The consumer runs as a StatefulSet
  with the ShadowPod strategy and ExchangeFence identity swap
  (Section~\ref{sec:identity-swap}). After replay, the operator re-checkpoints
  the shadow pod, creates a StatefulSet-owned replacement, and uses
  ExchangeFence for zero-loss traffic handoff. Checkpoints are transferred via
  the registry using the agent-assisted fast path (ms2m-agent).
\end{enumerate}

The comparison between configurations 1 and 2 isolates the effect of the
ShadowPod strategy on the same workload type (StatefulSet). The comparison
between configurations 2 and 3 reveals the effect of workload type (StatefulSet
vs.\ Deployment). The comparison between configurations 2 and 4 isolates the
effect of identity swap (and its associated agent-assisted transfer path) on the same workload type.

\subsubsection{Parameters}

Each configuration is evaluated at seven message rates: 10, 20, 40, 60, 80,
100, and 120 messages per second. Each rate--configuration combination is
repeated 10 times, yielding $4 \times 7 \times 10 = 280$ total migration runs.
The replay cutoff is set to 120 seconds. Between runs, message queues are purged
to prevent accumulation, checkpoint images are cleaned from worker nodes to
prevent disk exhaustion, and the consumer pod is verified to be ready and
processing messages before initiating the next migration.

\subsubsection{Downtime Measurement}
\label{sec:downtime-method}

Service downtime is measured using an external HTTP probe pod that sends
requests to the consumer's Service endpoint at 10\,ms intervals. Downtime is
the \textit{longest contiguous streak} of failed probes during the migration
window, with a 3\,s gap threshold to avoid merging unrelated failures. Total
migration time is the wall-clock duration from CR creation to Completed phase.
%% CONDENSED: expanded downtime measurement methodology
% A dedicated probe pod, deployed on a worker node separate from the consumer,
% continuously sends HTTP requests to the consumer's Service endpoint
% (consumer:8080) at 10ms intervals throughout each migration run. Each probe
% records a timestamp and result (UP or DOWN) to a log file.
% A streak is defined as a sequence of DOWN probes where consecutive failures
% are separated by no more than 3 seconds---this threshold accommodates the
% probe interval without merging unrelated failure events. This streak-based
% metric captures the actual service unavailability as perceived by clients,
% rather than the sum of scattered transient failures that may be caused by
% kube-proxy endpoint propagation delays.
% Although the consumer is primarily a message processor, the HTTP endpoint
% serves a dual purpose: it is the Kubernetes readiness probe target, and it
% provides synchronous access to the service's current processing state (e.g.,
% message count, last sequence number). Service availability as observed through
% HTTP reflects whether clients---including monitoring systems, load balancers,
% and dependent services---can reach the consumer. A period of HTTP
% unavailability therefore represents genuine service disruption, not merely a
% pause in message consumption.
% Total migration time is measured as the wall-clock duration from CR creation
% to the Completed phase, independent of the sum of per-phase durations recorded
% by the operator (which may exclude inter-phase reconciliation overhead).

\subsection{Results}
\label{sec:results}

\subsubsection{Total Migration Time}

Table~\ref{tab:total-time} presents the median total wall-clock migration time
($n=10$ per cell) for each configuration across all seven message rates.
Figure~\ref{fig:total-time} visualizes the trend.

%% CONDENSED: expanded total migration time analysis (two paragraphs)
% At low message rates (10\,msg/s), the ShadowPod configurations complete
% migration in 12.4--13.8\,s---a 73--76\% reduction from the Sequential baseline
% of 50.8\,s. This improvement stems almost entirely from eliminating the
% ${\sim}$38.5\,s restore phase, in which the Sequential strategy must scale down
% the StatefulSet, wait for pod termination, and create a new pod from the
% checkpoint image. At moderate rates (20--40\,msg/s), the ShadowPod
% configurations maintain sub-20\,s migration times while the Sequential baseline
% grows to 59--81\,s as replay time increases.
%
% At high message rates ($\geq$\,80\,msg/s), the replay phase dominates total
% migration time: the 120\,s replay cutoff accounts for over 73\% of the total
% duration across all configurations. The ShadowPod configurations still achieve
% a 22\% reduction (129.0--129.2\,s vs.\ 164.8\,s at 120\,msg/s), but the
% marginal benefit over Sequential is smaller because the replay duration is
% rate-dependent rather than strategy-dependent. At the intermediate rate of
% 60\,msg/s, the ShadowPod advantage remains substantial: SS-Shadow completes
% in 36.0\,s and D-Reg in 58.2\,s, compared to 157.5\,s for Sequential.
At low rates (10\,msg/s), ShadowPod configurations complete in
12.4--13.8\,s, a 73--76\% reduction from the Sequential baseline of 50.8\,s,
primarily by eliminating the 38.5\,s median restore phase. At high rates
($\geq$\,80\,msg/s), the 120\,s replay cutoff dominates, narrowing the
ShadowPod advantage to 22\%. At the intermediate rate of 60\,msg/s, SS-Shadow
completes in 36.0\,s vs.\ 157.5\,s for Sequential (77\% reduction).

The SS-Swap configuration shows similar total times to SS-Shadow at low rates
(15.8\,s vs.\ 13.8\,s at 10\,msg/s), with the additional 2\,s attributable to
the identity swap finalization. At high rates, SS-Swap is 10\,s slower than
SS-Shadow (141.6\,s vs.\ 129.0\,s at 120\,msg/s) due to the 17.5\,s finalize
phase, but still 14\% faster than Sequential.

\begin{figure}[t]
\centering
\begin{tikzpicture}
\begin{axis}[
  width=\columnwidth,
  height=6cm,
  ylabel={Total Migration Time (s)},
  xlabel={Message Rate (msg/s)},
  xmin=0, xmax=130,
  ymin=0, ymax=180,
  xtick={10,20,40,60,80,100,120},
  ytick={0,30,60,90,120,150,180},
  grid=major,
  grid style={gray!30},
  legend style={at={(0.02,0.98)}, anchor=north west, font=\scriptsize,
    legend columns=1},
  mark size=2pt,
  thick,
]
\addplot[mark=square*, blue] coordinates {
  (10,50.8) (20,59.2) (40,81.1) (60,157.5) (80,164.7) (100,165.0) (120,164.8)};
\addplot[mark=triangle*, red] coordinates {
  (10,13.8) (20,14.3) (40,19.5) (60,36.0) (80,117.2) (100,129.8) (120,129.0)};
\addplot[mark=diamond*, orange] coordinates {
  (10,15.8) (20,17.6) (40,23.1) (60,31.8) (80,58.3) (100,139.9) (120,141.6)};
\addplot[mark=o, green!50!black] coordinates {
  (10,12.4) (20,14.0) (40,20.3) (60,58.2) (80,126.0) (100,128.9) (120,129.2)};
\addplot[dashed, gray!60, forget plot] coordinates {(0,120) (130,120)};
\node[font=\tiny, gray] at (axis cs:65,125) {replay cutoff (120\,s)};
\legend{SS-Seq, SS-Shadow, SS-Swap, D-Reg}
\end{axis}
\end{tikzpicture}
\caption{Total wall-clock migration time (median, $n=10$) across message rates.
At low rates ($\leq$\,40\,msg/s), ShadowPod reduces time by 73--76\%. At high
rates where the 120\,s replay cutoff dominates, the reduction narrows to 22\%.
The dashed line marks the replay cutoff boundary.}
\label{fig:total-time}
\end{figure}

\begin{table*}[t]
\centering
\caption{Total Migration Time, median (seconds), $n=10$ per cell. Values
in parentheses indicate the interquartile range (Q1--Q3).}
\label{tab:total-time}
\begin{tabular}{lcccc}
\toprule
\textbf{Rate} & \textbf{SS-Seq} & \textbf{SS-Shadow} & \textbf{SS-Swap} & \textbf{D-Reg} \\
\midrule
10  & 50.8\,\scriptsize(50.3--80.1) & 13.8\,\scriptsize(12.3--13.8) & 15.8\,\scriptsize(14.9--17.4) & 12.4\,\scriptsize(12.2--13.3) \\
20  & 59.2\,\scriptsize(58.2--60.0) & 14.3\,\scriptsize(13.3--44.3) & 17.6\,\scriptsize(16.7--20.4) & 14.0\,\scriptsize(13.0--16.3) \\
40  & 81.1\,\scriptsize(80.2--105.9) & 19.5\,\scriptsize(18.7--84.8) & 23.1\,\scriptsize(21.9--28.4) & 20.3\,\scriptsize(18.9--29.4) \\
60  & 157.5\,\scriptsize(130.1--164.5) & 36.0\,\scriptsize(29.1--100.5) & 31.8\,\scriptsize(30.0--32.8) & 58.2\,\scriptsize(27.9--108.3) \\
80  & 164.7\,\scriptsize(163.9--165.2) & 117.2\,\scriptsize(74.7--129.3) & 58.3\,\scriptsize(52.5--67.5) & 126.0\,\scriptsize(103.4--129.7) \\
100 & 165.0\,\scriptsize(164.5--165.9) & 129.8\,\scriptsize(128.9--130.9) & 139.9\,\scriptsize(138.9--140.2) & 128.9\,\scriptsize(128.4--130.0) \\
120 & 164.8\,\scriptsize(164.3--165.5) & 129.0\,\scriptsize(128.5--130.1) & 141.6\,\scriptsize(141.1--142.0) & 129.2\,\scriptsize(128.9--131.4) \\
\bottomrule
\end{tabular}
\end{table*}


\subsubsection{Phase-by-Phase Breakdown}

Tables~\ref{tab:phases-10} and~\ref{tab:phases-60} present the per-phase
durations at 10 and 60\,msg/s respectively, illustrating the source of the
performance improvement at both ends of the rate spectrum.

\textbf{Checkpointing} is consistent at a median of 0.34\,s across all
configurations. \textbf{Transferring} takes 5.0--5.9\,s (I/O-bound registry
transfer). The \textbf{Restoring} phase shows the primary ShadowPod benefit:
Sequential requires a median of 38.5\,s (StatefulSet identity constraint), while
ShadowPod requires only 2.3--3.2\,s, a \textbf{92\% reduction}. The
\textbf{Replaying} phase scales with message rate: at 60\,msg/s, SS-Shadow
replays in 27.7\,s vs.\ 112.8\,s for Sequential, because fewer messages
accumulate during the shorter restore window. At $\geq$\,80\,msg/s, all
configurations hit the 120\,s cutoff. \textbf{Finalizing} is near-instantaneous
($<$\,0.1\,s). The SS-Swap configuration shows near-zero transfer time (0.19\,s at 60\,msg/s) due to the agent-assisted registry push (which avoids Kubernetes Job scheduling overhead), but a much longer Finalizing phase (14.73\,s) due to the identity swap procedure: re-checkpointing, pod replacement, and ExchangeFence queue draining. Figure~\ref{fig:phase-breakdown} visualizes the breakdown at
60\,msg/s.
%% CONDENSED: detailed per-phase prose (data in Tables 3-4)
% The \textbf{Checkpointing} phase is consistent across all configurations and
% rates at 0.31--0.45\,s (median ${\sim}$0.34\,s), as the same CRIU checkpoint
% operation is performed regardless of migration strategy. The checkpoint
% archive is approximately 27\,MB for the consumer workload.
%
% The \textbf{Transferring} phase takes 5.0--5.9\,s for all configurations, as
% all three use the in-cluster registry for checkpoint image transfer. This phase
% involves building an OCI image from the checkpoint archive, pushing it to the
% registry, and pulling it on the target node. The duration is remarkably stable
% across rates, confirming that it is I/O-bound rather than CPU-bound.
%
% The \textbf{Restoring} phase reveals the primary benefit of the ShadowPod
% strategy. The Sequential configuration consistently requires ${\sim}$38.5\,s
% across all rates because the StatefulSet must be scaled to zero, the source pod
% terminated, and a new pod created with the checkpoint image---all constrained
% by Kubernetes' StatefulSet identity enforcement. The ShadowPod configurations
% require only 2.3--3.2\,s, as the shadow pod is created as an independent pod
% alongside the still-running source. This represents a \textbf{92\% reduction}
% in restore duration ($38.4 \to 2.9$\,s at 60\,msg/s).
%
% The \textbf{Replaying} phase scales with message rate. At 10\,msg/s, replay
% completes in 4--7\,s across all configurations. At 60\,msg/s, the Sequential
% configuration's median replay time reaches 112.8\,s (near the 120\,s cutoff),
% while SS-Shadow shows 27.7\,s and D-Reg shows 49.3\,s---substantially lower
% because the source pod continues processing messages during the shorter restore
% window, resulting in fewer messages buffered in the replay queue. At rates
% $\geq$\,80\,msg/s, all configurations hit the 120\,s cutoff.
%
% The \textbf{Finalizing} phase is near-instantaneous for all configurations
% ($<$\,0.1\,s), consisting of control message exchange and resource cleanup.
%
% Figure~\ref{fig:phase-breakdown} visualizes the phase breakdown at 60\,msg/s,
% highlighting that the restore phase---which dominates the Sequential
% configuration---is nearly eliminated by the ShadowPod strategy.

\begin{figure}[t]
\centering
\begin{tikzpicture}
\begin{axis}[
  xbar stacked,
  width=\columnwidth,
  height=5.5cm,
  bar width=10pt,
  xlabel={Duration (s)},
  symbolic y coords={D-Reg, SS-Swap, SS-Shadow, SS-Seq},
  ytick=data,
  yticklabel style={font=\footnotesize},
  xmin=0, xmax=180,
  legend style={at={(0.5,-0.3)}, anchor=north, font=\scriptsize,
    legend columns=3},
  reverse legend,
  enlarge y limits=0.25,
]
% Stacked: Checkpoint, Transfer, Restore, Replay, Finalize (medians at 60 msg/s, n=10)
\addplot coordinates {(0.34,SS-Seq) (0.34,SS-Shadow) (0.39,SS-Swap) (0.33,D-Reg)};
\addplot coordinates {(5.35,SS-Seq) (5.64,SS-Shadow) (0.19,SS-Swap) (5.38,D-Reg)};
\addplot coordinates {(38.41,SS-Seq) (2.89,SS-Shadow) (3.02,SS-Swap) (2.50,D-Reg)};
\addplot coordinates {(112.80,SS-Seq) (27.68,SS-Shadow) (13.44,SS-Swap) (49.33,D-Reg)};
\addplot coordinates {(0.00,SS-Seq) (0.02,SS-Shadow) (14.73,SS-Swap) (0.04,D-Reg)};
\legend{Checkpoint, Transfer, Restore, Replay, Finalize}
\end{axis}
\end{tikzpicture}
\caption{Phase duration breakdown at 60\,msg/s (medians, $n=10$). The
Sequential configuration is dominated by the 38.4\,s restore phase. ShadowPod
reduces restore to 2.5--2.9\,s, shifting the bottleneck entirely to replay.}
\label{fig:phase-breakdown}
\end{figure}

\begin{table}[t]
\centering
\caption{Phase Durations (median seconds, $n=10$) at 10\,msg/s}
\label{tab:phases-10}
\begin{tabular}{lcccc}
\toprule
\textbf{Phase} & \textbf{SS-Seq} & \textbf{SS-Shadow} & \textbf{SS-Swap} & \textbf{D-Reg} \\
\midrule
Checkpointing  & 0.33  & 0.33  & 0.37 & 0.34 \\
Transferring   & 5.30  & 5.60  & 0.20 & 5.26 \\
Restoring      & 38.49 & 3.23  & 2.61 & 2.28 \\
Replaying      & 6.53  & 4.08  & 3.78 & 4.30 \\
Finalizing     & 0.01  & 0.02  & 8.64 & 0.03 \\
\midrule
\textbf{Total} & 50.8  & 13.8  & 15.8 & 12.4 \\
\bottomrule
\end{tabular}
\end{table}

\begin{table}[t]
\centering
\caption{Phase Durations (median seconds, $n=10$) at 60\,msg/s}
\label{tab:phases-60}
\begin{tabular}{lcccc}
\toprule
\textbf{Phase} & \textbf{SS-Seq} & \textbf{SS-Shadow} & \textbf{SS-Swap} & \textbf{D-Reg} \\
\midrule
Checkpointing  & 0.34  & 0.34  & 0.39 & 0.33 \\
Transferring   & 5.35  & 5.64  & 0.19 & 5.38 \\
Restoring      & 38.41 & 2.89  & 3.02 & 2.50 \\
Replaying      & 112.80 & 27.68 & 13.44 & 49.33 \\
Finalizing     & 0.00  & 0.02  & 14.73 & 0.04 \\
\midrule
\textbf{Total} & 157.5 & 36.0  & 31.8 & 58.2 \\
\bottomrule
\end{tabular}
\end{table}


\subsubsection{Service Downtime}

Table~\ref{tab:downtime} reports the measured service downtime (median, $n=10$)
using the HTTP probe methodology described in
Section~\ref{sec:downtime-method}.

The Sequential baseline exhibits a consistent median of 31.1\,s of downtime across
all rates (rate-independent, determined by the StatefulSet scale-down/up cycle).
All three ShadowPod configurations achieve \textbf{zero measured downtime}: D-Reg
recorded 0\,ms in all 70 runs; SS-Shadow recorded 0\,ms in 60/70 runs, with the
remaining 10 showing exactly 10\,ms (a single probe failure due to kube-proxy
endpoint propagation). SS-Swap recorded 0\,ms in 60 of 70 runs, with the remaining 10 runs at 20\,msg/s showing 10\,ms. This holds across all seven message rates.
%% CONDENSED: expanded downtime analysis
% Individual Sequential values range from 29.8s to 34.9s across all 70 runs.
% The zero-downtime result follows directly from the ShadowPod design: the
% source pod continues serving traffic via the Kubernetes Service throughout the
% entire migration, while the shadow pod becomes an additional Service endpoint
% upon readiness. During finalization, the source pod is removed (via StatefulSet
% scale-down or direct deletion), but the shadow pod is already receiving and
% serving requests---making the traffic handoff imperceptible to clients.

%% CONDENSED: downtime bar chart removed (3 of 4 configs are zero, figure adds no information)
% \begin{figure}[t]
% \centering
% ... downtime bar chart ...
% \caption{Measured service downtime...}
% \label{fig:downtime}
% \end{figure}

\begin{table}[t]
\centering
\caption{Service Downtime (median) by Configuration and Message Rate ($n=10$
per cell). SS-Seq values in seconds; ShadowPod values in milliseconds.}
\label{tab:downtime}
\begin{tabular}{lcccc}
\toprule
\textbf{Rate (msg/s)} & \textbf{SS-Seq} & \textbf{SS-Shadow} & \textbf{SS-Swap} & \textbf{D-Reg} \\
\midrule
10  & 31.2\,s & 0\,ms & 0\,ms & 0\,ms \\
20  & 31.2\,s & 10\,ms & 10\,ms & 0\,ms \\
40  & 31.2\,s & 0\,ms & 0\,ms & 0\,ms \\
60  & 31.2\,s & 0\,ms & 0\,ms & 0\,ms \\
80  & 31.1\,s & 0\,ms & 0\,ms & 0\,ms \\
100 & 31.1\,s & 0\,ms & 0\,ms & 0\,ms \\
120 & 30.9\,s & 0\,ms & 0\,ms & 0\,ms \\
\bottomrule
\end{tabular}
\end{table}

\subsubsection{Message Loss}

Across all 280 migration runs, no messages were lost during migration. The
message replay mechanism successfully synchronized state between source and
target in every completed run across all four configurations and all seven
message rates, confirming the zero-loss guarantee of the MS2M framework.


\subsection{Discussion}
\label{sec:discussion}

\subsubsection{Interpretation}

%% CONDENSED: original "Key Findings" subsubsection that restated results
% \textbf{SHADOW eliminates the restore bottleneck and service downtime.}
% Restore duration decreases from 38.4\,s to 2.9\,s (92\% reduction), and
% downtime drops from ${\sim}$31\,s to 0\,ms across all 70 ShadowPod runs on
% StatefulSet workloads. Total migration time decreases by 73\% at low rates
% (50.8\,s $\to$ 13.8\,s at 10\,msg/s).
%% CONDENSED: expanded first key finding
% The comparison between statefulset-sequential and statefulset-shadowpod isolates
% the effect of the ShadowPod strategy on the same workload type. [... full text
% preserved in earlier CONDENSED block ...]

The comparison between configurations~1 and~2 isolates the effect of the
ShadowPod strategy on the same workload type (StatefulSet). The 92\% reduction
in restore duration confirms that the StatefulSet identity constraint, which
prior work treated as an inherent limitation, can be circumvented by
decoupling traffic routing (label selectors) from pod ownership (controller
references). Kubernetes Services and StatefulSet identity operate on
independent mechanisms, and this separation is what makes zero-downtime
migration possible without requiring a workload type change.

The comparison between configurations~2 and~3 reveals that the ShadowPod
strategy performs similarly on both StatefulSet and Deployment workloads,
confirming that the approach is not tied to a specific workload controller.
The small differences in replay duration (27.7\,s vs.\ 49.3\,s at 60\,msg/s)
are attributable to scheduling variance rather than a systematic effect of
workload type.

The comparison between configurations~2 and~4 isolates the effect of
identity swap. SS-Swap's transfer phase is 25x faster than
SS-Shadow (0.19\,s vs.\ 5.64\,s at 60\,msg/s) because the agent-assisted
registry push avoids Kubernetes Job scheduling overhead, but its finalize phase is 700x longer (14.73\,s vs.\ 0.02\,s) due to
the re-checkpoint, pod replacement, and ExchangeFence procedure. The net effect
is a modest increase in total migration time (31.8\,s vs.\ 36.0\,s at 60\,msg/s
for SS-Swap vs.\ SS-Shadow), but with the benefit of restoring full StatefulSet
ownership of the migrated workload.

%% CONDENSED: expanded replay analysis
% At intermediate rates (40--60\,msg/s), the ShadowPod advantage remains
% substantial due to shorter replay queues: at 60\,msg/s, SS-Shadow completes
% in 36.0\,s vs.\ 157.5\,s for Sequential (77\% reduction).
% This behavior reflects an inherent constraint: the replay mechanism requires
% the target service's message processing rate to exceed the incoming message
% rate for the queue to drain before the cutoff fires. When
% $\mu_{\text{target}} < \lambda$, the replay queue grows unboundedly and the
% cutoff triggers with messages still pending. In this case, the shadow pod's
% state after migration is partially synchronized---it has replayed all messages
% up to the cutoff point, but messages arriving during the final window are
% processed only after migration completes. The evaluation confirms this:
% at $\geq$\,80\,msg/s (approaching the consumer's ${\sim}$65\,msg/s effective
% throughput ceiling), replay consistently reaches the 120\,s cutoff.
At high message rates ($\geq$\,80\,msg/s), the replay cutoff dominates total
migration time across all configurations. This reveals a limitation of
the MS2M replay mechanism: when $\mu_{\text{target}} < \lambda$, the replay
queue grows faster than it can be drained, and the cutoff fires with messages
still pending. The ShadowPod strategy does \textit{reduce} the replay
queue length compared to Sequential (because the source continues processing
during the shorter restore window), but cannot eliminate the rate-dependent
bottleneck. Techniques such as batched message processing, increased consumer
parallelism, or CRIU pre-dump for incremental checkpointing would address this
regime.

%% CONDENSED: expanded zero message loss finding
% \textbf{Zero message loss across all 210 runs.} SHADOW's message replay
% mechanism maintained zero message loss across all configurations, workload
% types, and message rates, validating the correctness of the ShadowPod strategy
% for both StatefulSet-managed and Deployment-managed workloads.
Zero message loss across all 280 runs confirms the correctness of the
concurrent source-shadow operation: the fanout exchange ensures that both
pods receive identical message streams on separate queues, and the control
message protocol guarantees an orderly handoff.

\subsubsection{Comparison with Prior Work}

%% CONDENSED: expanded comparison with prior work
% The evaluation in~\cite{dinhtuan2025k8s} was conducted on Google Compute Engine
% e2-medium VMs (2 vCPUs, 4\,GB RAM) with a Java/Spring Boot consumer and a
% Python migration manager. The present evaluation uses bare-metal servers
% (4 vCPUs, 8\,GB RAM) with a Python consumer and a Go-based Kubernetes operator.
% These differences affect absolute timing values but do not invalidate the
% relative comparison between configurations, as all three share the same
% infrastructure and workload.
%
% Notably, the checkpoint and transfer phase durations are comparable between
% the two environments (checkpoint ${\sim}$0.34\,s vs.\ ${\sim}$0.4\,s in prior
% work; transfer ${\sim}$5.4\,s vs.\ ${\sim}$6\,s), suggesting that these phases
% are dominated by I/O operations (CRIU dump, OCI image construction, registry
% push/pull) rather than compute capacity. The restore phase for Sequential
% migration is also comparable (${\sim}$38.5\,s vs.\ ${\sim}$39\,s), confirming
% that this bottleneck is inherent to the StatefulSet identity constraint rather
% than hardware-dependent.
The prior evaluation~\cite{dinhtuan2025k8s} used GCE e2-medium VMs with a
Java consumer. Despite different infrastructure, checkpoint (median 0.34\,s
vs.\ 0.4\,s), transfer (median 5.4\,s vs.\ 6\,s), and
Sequential restore (median 38.5\,s vs.\ 39\,s) durations are
comparable, confirming these phases are I/O-bound and the restore bottleneck is
intrinsic to StatefulSet identity constraints.

\subsubsection{Trade-offs and Operator Overhead}

The ShadowPod strategy for StatefulSets transitions the workload to a standalone
shadow pod; the Sequential strategy remains available for workloads requiring
StatefulSet guarantees. The direct transfer mechanism (not evaluated in this paper) can
eliminate registry dependency at the cost of deploying the ms2m-agent DaemonSet
with \texttt{hostPath} mounts; the agent-assisted registry push used in
SS-Swap provides a partial benefit without bypassing the registry. The operator's reconcile loop overhead (API
round-trips per phase transition) is negligible relative to phase durations and
is justified by automatic retry, declarative lifecycle management, and native
observability.
%% CONDENSED: expanded trade-offs and operator overhead (two subsubsections in original)
% \subsubsection{Trade-offs}
% The ShadowPod strategy for StatefulSet workloads (statefulset-shadowpod)
% introduces a structural change: after migration, the workload transitions from
% StatefulSet management to a standalone shadow pod. For event-driven
% microservices whose routing is managed by the message broker, this trade-off
% is acceptable. For workloads that require StatefulSet guarantees (ordered
% scaling, stable storage), the Sequential strategy remains available---though
% with the associated $\sim$31\,s downtime.
%
% The direct transfer mechanism (not included in the main evaluation matrix
% to isolate the ShadowPod effect) eliminates the registry dependency at the
% cost of deploying the ms2m-agent DaemonSet on all worker nodes. The agent's
% resource footprint is minimal, but the DaemonSet requires \texttt{hostPath}
% volume mounts that must be permitted by cluster security policies.
%
% \subsubsection{Operator Overhead}
% The SHADOW operator introduces reconcile loop overhead compared to a direct
% script-based approach. Each phase transition involves a Kubernetes API server
% round-trip to update the resource status. However, the phase chaining
% optimization (Section~\ref{sec:statemachine}) reduces this overhead by executing
% synchronously-completed phases within a single API call. The remaining overhead
% is negligible relative to the phase durations (milliseconds vs.\ seconds) and is
% justified by the operational benefits: automatic retry on transient failures,
% declarative resource lifecycle, and integration with Kubernetes' native
% observability (events, status conditions, audit logs).


%% ============================================================================
\section{Related Work}
\label{sec:related}
%% ============================================================================

%% CONDENSED: expanded container live migration paragraph
% \textbf{Container Live Migration.}
% Live migration of containers has been surveyed comprehensively by Soussi
% et al.~\cite{soussi2024survey}, who categorize techniques along three
% dimensions: optimization method (pre-copy, post-copy, hybrid), traffic
% relocation strategy, and deployment context (cloud, edge, MEC). All
% memory-centric approaches face a fundamental trade-off: pre-copy migration
% iteratively transfers dirty memory pages while the service runs, reducing
% downtime but increasing total migration time proportionally to the dirty page
% rate~\cite{lu2023mbdpc}; post-copy migration resumes the service immediately
% and fetches pages on demand, achieving low downtime at the cost of degraded
% post-migration performance~\cite{tazzioli2023postcopy}. SHADOW sidesteps this
% trade-off entirely by reconstructing state through application-level message
% replay rather than memory-level synchronization.
\textbf{Container Live Migration.}
Live migration of containers has been surveyed comprehensively by Soussi
et al.~\cite{soussi2024survey}, who categorize techniques along three
dimensions: optimization method (pre-copy, post-copy, hybrid), traffic
relocation strategy, and deployment context (cloud, edge, MEC). All
memory-centric approaches face a fundamental trade-off: pre-copy migration
iteratively transfers dirty memory pages while the service runs, reducing
downtime but increasing total migration time proportionally to the dirty page
rate~\cite{lu2023mbdpc}; post-copy migration resumes the service immediately
and fetches pages on demand, achieving low downtime at the cost of degraded
post-migration performance~\cite{tazzioli2023postcopy}. SHADOW avoids this
trade-off entirely by reconstructing state through application-level message
replay rather than memory-level synchronization.

%% CONDENSED: expanded CRIU-based pod migration paragraph
% \textbf{CRIU-based Pod Migration in Kubernetes.}
% Forensic Container Checkpointing (FCC)~\cite{k8sfcc} provides the kubelet API
% for CRIU-based container checkpointing, forming the foundation for checkpoint
% creation in both SHADOW and related systems. KubeSPT~\cite{kubespt2025}
% addresses stateful pod migration in Kubernetes through three mechanisms: a
% T-Proxy component that freezes the pod's network namespace to maintain TCP
% connections during migration, a Hot Data and Lazy-Restore method that selectively
% restores frequently accessed memory pages first, and a decoupled pod recreation
% procedure that parallelizes Kubernetes pod initialization with checkpoint
% transfer. KubeSPT reports 86--93\% downtime reduction for memory-intensive
% workloads such as Redis. However, KubeSPT focuses on preserving raw memory state
% and network connections---it does not address application-level state consistency
% for event-driven microservices. SHADOW takes a complementary approach:
% rather than optimizing memory page transfer, it leverages the application's
% messaging infrastructure to guarantee both state consistency and zero message
% loss. Furthermore, SHADOW's ShadowPod strategy enables true zero-downtime
% migration by keeping the source pod serving traffic throughout the entire
% process, whereas KubeSPT's downtime is bounded by checkpoint transfer and
% restore duration.
\textbf{CRIU-based Pod Migration in Kubernetes.}
FCC~\cite{k8sfcc} provides the kubelet API for CRIU checkpointing.
KubeSPT~\cite{kubespt2025} addresses stateful pod migration through T-Proxy
(TCP connection preservation), lazy-restore of hot memory pages, and decoupled
pod recreation, reporting 86--93\% downtime reduction for memory-intensive
workloads. However, KubeSPT focuses on raw memory state, not application-level
consistency for event-driven services. SHADOW instead uses messaging
infrastructure for state consistency and zero message loss, with the ShadowPod
strategy achieving true zero-downtime by keeping the source serving throughout.
%% CONDENSED: other CRIU-based approaches
% Tazzioli et al.~\cite{tazzioli2023postcopy} integrated post-copy live
% migration into Kubernetes, achieving 65\% downtime reduction by lazily
% fetching memory pages on demand. The Ubiquitous Migration Solution
% (UMS)~\cite{ums} introduces a three-phase migration process with a Frontman
% container for traffic management during migration. Ma et al.~\cite{ma2019live}
% proposed a Docker-specific method for edge environments that transfers only
% the writable filesystem layer, reducing migration time by 56--80\%.
Tazzioli et al.~\cite{tazzioli2023postcopy} integrated post-copy live migration
into Kubernetes, achieving 65\% downtime reduction by lazily fetching memory
pages on demand. The Ubiquitous Migration Solution (UMS)~\cite{ums} introduces
a three-phase migration process with a Frontman container for traffic management
during migration. Ma et al.~\cite{ma2019live} proposed a Docker-specific method
for edge environments that transfers only the writable filesystem layer,
reducing migration time by 56--80\%. None of these approaches support concurrent
source-target operation during state synchronization, which is what makes
SHADOW's zero-downtime guarantee possible.

\textbf{Stateful Microservice Migration.}
Calagna et al.~\cite{calagna2024coat} address stateful migration at the edge
with two contributions: COAT (Container OverlAy TCP), a network solution based
on Open vSwitch that preserves established TCP connections during migration by
reconstructing the network namespace at the destination, and PAM
(Processing-Aware Migration), an analytical model that predicts migration KPIs
accounting for container runtime overhead. However, COAT operates on Podman with
imperative orchestration scripts rather than Kubernetes-native mechanisms and
focuses on connection-oriented edge workloads rather than event-driven
microservices. Laigner et al.~\cite{laigner2025transactional} survey the state
management landscape for transactional cloud applications, identifying that
microservice frameworks typically embed state within the application runtime and
rely on message queues for inter-service communication. This is precisely the
property that SHADOW uses for state reconstruction through message replay.
%% CONDENSED: expanded stateful microservice migration paragraphs
% \textbf{Stateful Microservice Migration.}
% Calagna et al.~\cite{calagna2024coat} address stateful migration at the edge
% with two contributions: COAT (Container OverlAy TCP), a network solution based
% on Open vSwitch that preserves established TCP connections during migration by
% reconstructing the network namespace at the destination, and PAM
% (Processing-Aware Migration), an analytical model that predicts migration KPIs
% accounting for container runtime overhead. Their work demonstrates the
% importance of modeling the full migration stack---from CRIU internals through
% container runtime layers---for accurate performance prediction. However, COAT
% operates on Podman with imperative orchestration scripts rather than
% Kubernetes-native mechanisms and focuses on connection-oriented edge workloads
% (MQTT broker, Memcached) rather than event-driven microservices.
%
% Laigner et al.~\cite{laigner2025transactional} survey the state management
% landscape for transactional cloud applications, identifying that microservice
% frameworks typically embed state within the application runtime and rely on
% message queues for inter-service communication. This architectural pattern---where
% service state is causally derived from processed messages---is precisely the
% property that MS2M and SHADOW exploit for state reconstruction through message
% replay, rather than requiring direct access to in-memory state.

%% CONDENSED: expanded Kubernetes Operators paragraph
% \textbf{Kubernetes Operators for Lifecycle Management.}
% The Operator pattern~\cite{operatorpattern} has been widely adopted for managing
% complex application lifecycles in Kubernetes, including database operators
% (e.g., for PostgreSQL, MySQL) and stateful middleware. Both
% KubeSPT~\cite{kubespt2025} and SHADOW use Custom Resource Definitions to drive
% migration as a Kubernetes-native workflow. SHADOW extends this pattern with an
% idempotent state machine reconciler that supports multiple migration strategies
% (Sequential, ShadowPod) and transfer modes (Registry, Direct), enabling
% declarative migration orchestration within Kubernetes' desired-state model.
\textbf{Kubernetes Operators for Lifecycle Management.}
The Operator pattern~\cite{operatorpattern} has been widely adopted for managing
complex application lifecycles in Kubernetes, including database operators
(e.g., for PostgreSQL, MySQL) and stateful middleware. Both
KubeSPT~\cite{kubespt2025} and SHADOW use Custom Resource Definitions to drive
migration as a Kubernetes-native workflow. SHADOW extends this pattern with an
idempotent state machine reconciler that supports multiple migration strategies
(Sequential, ShadowPod) and transfer modes (Registry, Direct), supporting
declarative migration orchestration within Kubernetes' desired-state model.


%% ============================================================================
\section{Conclusion}
\label{sec:conclusion}
%% ============================================================================

%% CONDENSED: original conclusion (restated abstract/results)
% This paper presented SHADOW, a Kubernetes-native framework for zero-downtime
% stateful microservice migration that builds on the MS2M message replay approach.
% SHADOW's ShadowPod strategy enables concurrent source-target operation by
% creating a shadow pod outside the workload controller's ownership, carrying the
% same application labels for Service-based traffic routing. We demonstrated that
% this strategy is applicable to both Deployment-managed and StatefulSet-managed
% workloads---the latter by scaling down the StatefulSet during finalization rather
% than requiring a sequential stop-recreate cycle. SHADOW's direct node-to-node
% transfer mechanism bypasses the container registry entirely through a lightweight
% DaemonSet agent.
%
% The evaluation on a bare-metal Kubernetes cluster across three configurations,
% seven message rates, and 210 total migration runs confirmed that SHADOW reduces
% restore phase duration by 92\% (from ${\sim}$38.4\,s to ${\sim}$2.9\,s) and
% eliminates service downtime entirely (from ${\sim}$31\,s to 0\,ms), while
% maintaining zero message loss across all runs. At low message rates
% (10\,msg/s), total migration time decreases by 73\% (from 50.8\,s to 13.8\,s).
% At high rates where the replay cutoff dominates, the improvement is 22\% at
% 120\,msg/s.
%
% SHADOW demonstrates that encoding migration procedures as a Kubernetes-native
% control loop provides practical benefits beyond performance: declarative resource
% management, automatic failure recovery, and built-in observability through
% standard Kubernetes mechanisms.
%
% Future work includes integrating CRIU pre-dump support for incremental
% checkpointing to reduce both checkpoint size and transfer time, extending
% migration support to multi-container pods, and investigating scheduler
% integration for automatic migration triggering based on cluster resource
% pressure or node maintenance events.
SHADOW shows that Kubernetes' StatefulSet identity constraint, previously
treated as an unavoidable source of migration downtime, can be circumvented at the
application level by separating traffic routing (label-based) from pod ownership
(controller-based). The ShadowPod strategy uses this separation for
concurrent source-target operation,
eliminating service downtime entirely while preserving zero message loss. This
result holds across both Deployment-managed and StatefulSet-managed workloads,
confirming that the approach generalizes beyond a single workload controller.

For StatefulSet workloads that require full controller ownership after migration,
the identity swap procedure with ExchangeFence re-checkpoints the shadow pod and
creates a StatefulSet-adopted replacement, adding 9--17\,s to finalization but
restoring ordered scaling and crash recovery guarantees.

Beyond the performance improvements (92\% restore reduction, 73\% total time
reduction at low rates), SHADOW confirms the practical value of encoding
migration procedures as a Kubernetes-native control loop. The idempotent state
machine reconciler provides automatic failure recovery, declarative lifecycle
management, and built-in observability; these are operational properties that are
difficult to achieve with external orchestration scripts. The evaluation across
280 migration runs with zero failures confirms the reliability of this approach
under varying message rates.

The remaining performance bottleneck is the replay phase at high message rates,
where the cutoff mechanism bounds migration time but leaves the shadow pod with
partially-synchronized state. Future work will address this through CRIU
pre-dump support for incremental checkpointing (reducing both checkpoint size
and transfer time), batched message replay to increase effective consumer
throughput, and evaluation of the direct node-to-node transfer
mode (which bypasses the registry entirely) across additional workload profiles. Additional directions include
scheduler integration for automatic migration triggering based on cluster
resource pressure, and extending support to multi-container pods.


%% ============================================================================
% References
%% ============================================================================

\balance
\bibliographystyle{IEEEtran}
\bibliography{references}

\end{document}